---
title: "Reproducible Research: Course Project 1"
author: "Rich Florez"
date: "September 11, 2016"
output: html_document
---
##Introduction##

This assignment makes use of data from a personal activity monitoring device. This device collects data at 5 minute intervals through out the day. The data consists of two months of data from an anonymous individual collected during the months of October and November, 2012 and include the number of steps taken in 5 minute intervals each day.

The data for this assignment can be downloaded from the course web site:

*Dataset:* [Activity monitoring data (52k)](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip)

The variables included in this dataset are:

- **steps:** Number of steps taking in a 5-minute interval (missing values are coded as 𝙽𝙰)
- **date:** The date on which the measurement was taken in YYYY-MM-DD format
- **interval:** Identifier for the 5-minute interval in which measurement was taken

The dataset is stored in a comma-separated-value (CSV) file and there are a total of 17,568 observations in this dataset.

##Loading and preprocessing the data##

The following packages/libraries were used in the analysis:

```{r}
```

Below is a summary of the dataset within the relevant working directory:

```{r}
setwd("~/datasciencecoursera/ReproRsrch/ProjectWk2/")
activity <- read.csv("activity.csv")
summary(activity)
```


##What is the mean total number of steps taken per day?##

*For this part of the assignment, we ignore the missing values in the dataset.*

- Calculate the total number of steps taken per day

```{r}
steps_per_day <- aggregate(steps ~ date, activity, sum)
head(steps_per_day)
```

- Histogram of the total number of steps taken each day

```{r}
hist(steps_per_day$steps,col="grey",main=("Total Steps Per Day"),xlab="Number of Steps")
```

- Calculate and report the mean and median of the total number of steps taken per day

```{r}
mean_steps <- mean(steps_per_day$steps)
median_steps <- median(steps_per_day$steps)
```

The mean is **`r format(round(mean_steps,2))`** and the median is **`r median_steps`**

##What is the average daily activity pattern?##

- Time series plot of the 5-minute interval and the average number of steps taken, averaged across all days: 

```{r}
steps_per_interval <- aggregate(steps ~ interval, activity, mean)
plot(steps_per_interval$interval,steps_per_interval$steps,type="l",main="Average Steps per Day per Interval",xlab="Interval", ylab="Number of Steps")
```

- Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
max_steps_per_interval <- steps_per_interval[which.max(steps_per_interval$steps),1]
val_max_steps_per_interval <- max(steps_per_interval$steps)
```

The 5-minute interval with the average maximum number of steps is **`r (max_steps_per_interval)`** at an average value of **`r format(round(val_max_steps_per_interval,2))`**

##Imputing missing values##

```{r}

```

##Are there differences in activity patterns between weekdays and weekends?##

```{r}

```
